---
title: "Aufgabe 2"
author: "windijon"
date: '2022-04-29'
output: html_document
---

### Task 1: Import your data

Load the necessary libraries.

```{r}

library(readr)        # to import tabular data (e.g. csv)
library(dplyr)        # to manipulate (tabular) data
library(ggplot2)      # to visualize data
library(sf)           # to handle spatial vector data
library(terra)        # To handle raster data
library(lubridate)    # To handle dates and times

```

Import the wild-boar data "wildschwein_BE_2056" as csv and create a Spatial Object.

```{r}

wildschwein_BE <- read_delim("wildschwein_BE_2056.csv",",")
View(wildschwein_BE)

wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056, remove = FALSE)
wildschwein_BE

```

### Task 2: Getting an overview

Calculate the time-lag between subsequent positions.

```{r}

wildschwein <- mutate(wildschwein_BE,timelag_s = as.integer(difftime(time1 = lead(DatetimeUTC), time2 = DatetimeUTC), units = "secs"))
wildschwein

```

How many individuals were tracked?

```{r}

wildschwein <- group_by(wildschwein,TierID)
wildschwein

nr_individual <- wildschwein %>% 
  group_keys()  %>% 
  nrow()
nr_individual

```

For how long were the individual tracked? Are there gaps? Were all individuals tracked concurrently or sequentially?

```{r}

ggplot(data = wildschwein, aes(x = DatetimeUTC, y = TierID)) +
  geom_line()

ggplot(data = wildschwein, aes(x = TierID, y = timelag_s)) +
  geom_boxplot()

# find and replace outliers with NA
wildschwein$timelag_s[wildschwein$timelag_s<=0]<-NA

ggplot(data = wildschwein, aes(x = TierID, y = timelag_s)) +
  geom_boxplot()

summary <- wildschwein %>%
  group_by(TierID) %>%
  summarise(         
    max = max(DatetimeUTC),
    min = min(DatetimeUTC),
    total_timelag = difftime(time1 = max, time2 = min),
    mean_timelag = mean(timelag_s, na.rm = T),
    sd_timelag = sd(timelag_s, na.rm = T),
    timelag_gabs = sum(is.na(timelag_s)))
summary

```

What is the temporal sampling interval between the locations?

```{r}

ggplot(data = wildschwein, aes(x = timelag_s)) +
  geom_histogram(bins = 100)

ggplot(data = wildschwein, aes(x = timelag_s)) +
  geom_histogram(bins = 200) + 
  scale_x_continuous(limits=c(0,4000))

```

### Task 3: Deriving movement parameters I: Speed

Calculate the Euclidean distance (step-length) and the speed.

```{r}

wildschwein <- wildschwein %>%
  mutate(steplength_m = sqrt((E-lead(E,1))^2 + (N-lead(N,1))^2),
         speed_m_s = steplength_m / timelag_s)

View(wildschwein)

```

### Task 4: Cross-scale movement analysis

Import the wild-boar data "Caro60" as csv. Reduce the granularity of our sampling interval by selecting every 3rd, 6th and 9th position.

```{r}

caro_st <- read_delim("caro60.csv",",")
View(caro_st)

caro <- st_as_sf(caro_st, coords = c("E", "N"), crs = 2056, remove = FALSE)
caro

seq_3 <- seq(from = 1, to = nrow(caro), by = 3)
caro_3 <- slice(.data = caro, n = seq_3)

seq_6 <- seq(from = 1, to = nrow(caro), by = 6)
caro_6 <- slice(.data = caro, n = seq_6)

seq_9 <- seq(from = 1, to = nrow(caro), by = 9)
caro_9 <- slice(.data = caro, n = seq_9)

nrow(caro)
nrow(caro_3)
nrow(caro_6)
nrow(caro_9)

```

Calculate time-lag, step-length and speed.

```{r}

caro_1 <- caro %>%
  mutate(timelag_s = as.integer(difftime(time1 = lead(DatetimeUTC), time2 = DatetimeUTC), units = "secs"),
         steplength_m = sqrt((E-lead(E,1))^2 + (N-lead(N,1))^2),
         speed_m_s = steplength_m / timelag_s,
         Trajectory = "1 minutes")
caro_1

caro_3 <- caro_3 %>%
  mutate(timelag_s = as.integer(difftime(time1 = lead(DatetimeUTC), time2 = DatetimeUTC), units = "secs"),
         steplength_m = sqrt((E-lead(E,1))^2 + (N-lead(N,1))^2),
         speed_m_s = steplength_m / timelag_s,
         Trajectory = "3 minutes")
caro_3

caro_6 <- caro_6 %>%
  mutate(timelag_s = as.integer(difftime(time1 = lead(DatetimeUTC), time2 = DatetimeUTC), units = "secs"),
         steplength_m = sqrt((E-lead(E,1))^2 + (N-lead(N,1))^2),
         speed_m_s = steplength_m / timelag_s,
         Trajectory = "6 minutes")
caro_6

caro_9 <- caro_9 %>%
  mutate(timelag_s = as.integer(difftime(time1 = lead(DatetimeUTC), time2 = DatetimeUTC), units = "secs"),
         steplength_m = sqrt((E-lead(E,1))^2 + (N-lead(N,1))^2),
         speed_m_s = steplength_m / timelag_s,
         Trajectory = "9 minutes")
caro_9

caro_merge <- rbind(caro_1, caro_3, caro_6, caro_9)

```

Compare the speeds visually in a line plot and visualize the trajectories in a map.

```{r}

Caro_speed_Cross_Scale <- ggplot(data = caro_merge) +
  geom_line(aes(x = DatetimeUTC, y = speed_m_s, color = Trajectory)) +
  labs(y = "Speed [m/s]", x = "Time") +
  ggtitle("Cross-Scale Movement Analysis") +
  theme_minimal()
Caro_speed_Cross_Scale

Caro_1_3 <- ggplot(data = subset(caro_merge, caro_merge$Trajectory == "1 minutes"), aes(x = E, y = N, color = Trajectory)) +
  geom_path(alpha = 0.4) +
  geom_point(alpha = 0.4) +
  geom_path(data = subset(caro_merge, caro_merge$Trajectory == "3 minutes"), alpha = 0.8) +
  geom_point(data = subset(caro_merge, caro_merge$Trajectory == "3 minutes"), alpha = 0.8) + 
  ggtitle("Comparing original- with 3 minutes-resampled data") +
  theme_minimal()
Caro_1_3

Caro_1_6 <- ggplot(data = subset(caro_merge, caro_merge$Trajectory == "1 minutes"), aes(x = E, y = N, color = Trajectory)) +
  geom_path(alpha = 0.4) +
  geom_point(alpha = 0.4) +
  geom_path(data = subset(caro_merge, caro_merge$Trajectory == "6 minutes"), alpha = 0.8) +
  geom_point(data = subset(caro_merge, caro_merge$Trajectory == "6 minutes"), alpha = 0.8) + 
  ggtitle("Comparing original- with 6 minutes-resampled data") +
  theme_minimal()
Caro_1_6

Caro_1_9 <- ggplot(data = subset(caro_merge, caro_merge$Trajectory == "1 minutes"), aes(x = E, y = N, color = Trajectory)) +
  geom_path(alpha = 0.4) +
  geom_point(alpha = 0.4) +
  geom_path(data = subset(caro_merge, caro_merge$Trajectory == "9 minutes"), alpha = 0.8) +
  geom_point(data = subset(caro_merge, caro_merge$Trajectory == "9 minutes"), alpha = 0.8) + 
  ggtitle("Comparing original- with 9 minutes-resampled data") +
  theme_minimal()
Caro_1_9

ggsave(plot = Caro_speed_Cross_Scale, filename = "Caro_Speed_Cross_Scale.pdf", width = 17, height = 14, units = "cm", dpi = 5000)
ggsave(plot = Caro_1_3, filename = "Caro_1_3.pdf", width = 17, height = 14, units = "cm", dpi = 5000)
ggsave(plot = Caro_1_6, filename = "Caro_1_6.pdf", width = 17, height = 14, units = "cm", dpi = 5000)
ggsave(plot = Caro_1_9, filename = "Caro_1_9.pdf", width = 17, height = 14, units = "cm", dpi = 5000)

```


### Task 5: Deriving movement parameters II: Rolling window functions

Install "zoo" package, load it into the session and use the function rollmean() to smooth the calculated speed

```{r}

library(zoo)

caro_1 <- caro_1 %>%
  mutate(rollmean = speed_m_s, 
         window_sizes_k = "1")

caro_3 <- caro_1 %>%
  mutate(rollmean = rollmean(speed_m_s, k = 3, fill = NA,align = "left"), 
         window_sizes_k = 3)


caro_6 <- caro_1 %>%
  mutate(rollmean = rollmean(speed_m_s, k = 6, fill = NA,align = "left"), 
         window_sizes_k = 6)

caro_9 <- caro_1 %>%
  mutate(rollmean = rollmean(speed_m_s, k = 9, fill = NA,align = "left"), 
         window_sizes_k = 9)

caro_merge <- rbind(caro_1, caro_3, caro_6, caro_9)

Caro_speed_moving_window <- ggplot(data = caro_merge) +
  geom_line(aes(x = DatetimeUTC, y = rollmean, color = window_sizes_k)) +
  labs(y = "Speed [m/s]", x = "Time") +
  ggtitle("Moving Window Movement Analysis") + labs(colour = "Moving Window Sizes [k]") +
  theme_minimal()
Caro_speed_moving_window

ggsave(plot = Caro_speed_moving_window, filename = "Caro_Speed_Moving_Window.pdf", width = 17, height = 14, units = "cm", dpi = 5000)

```

